Empirical Study of Scientific Computer Software: Objectives and Research Questions

### 2020-04-21 OverallResearchProposal/ObjectivesAndResearchQuestions.tex
update with comments

In the above definition, software is considered to be an application (app)
and/or a library.  The outcome of the development process is not just software.
It is \{software + artifacts\}.  \wss{softwartifacts?} The artifacts consist of
such work products as documentation, build scripts, code, test cases,
contributor's guides, etc.  The quality of sustainability cannot be assessed by
looking at the software alone, or the code alone; it depends on everything,
including the processes and methodologies used.  Further information on the
relevant definitions and context is given in the next section.

The notion of ``reasonable amount of time and energy'' is part of the objective,
but reasonable is not defined in the objective.  What is reasonable is going to
depend on the context.  Problems that are large, complex and/or critical are
going to have a greater ``budget'' of time and energy.  When comparing different
alternatives for processes and methodologies, preference would be given to the
alternative that costs less time and energy, assuming that this alternative
achieves the same level of sustainability as its competition.  The assumptions
listed in the next section outline the typical space of problems where it makes
sense to aim for sustainable \{software + artifacts\}.

@@

\begin{description}
\item[Knowledge] Facts, definitions, theories, assumptions, etc. \wss{needs a
    proper definition, with citation(s)}
\item[Process] Organization of the software development activities.  \wss{needs a
    proper definition, with citation(s)}
\item[Principles] Principles are ``general and abstract statements describing
  desirable properties of software processes and products.''  \citep[p.\
  41]{GhezziEtAl2003}
\item[Methods] ``Methods are general guidelines that govern the execution of some activity;
  they are rigorous, systematic, and disciplined approaches.''  \citep[p.\
  41]{GhezziEtAl2003}
\item[Techniques] ``Techniques are more technical and mechanical than methods.''
  \cite[p.\ 41]{GhezziEtAl2003}
\item[Methodology] Combination of methods and techniques.  \citep[p.\
  41]{GhezziEtAl2003}
\item[Tools] ``Tools ... are developed to support the application of techniques,
  method and methodologies.'' \cite[p.\ 41]{GhezziEtAl2003}

@@

This definition has brought in several other qualities.  These qualities are
defined in the ``Quality  Definitions of Qualities'' document.
\wss{Reproducibility above might be changed to replicability, or possibly both
  terms should be included?}
\wss{Should performance be included in the list of qualities?}

\item[Software] Application and/or library.
\item[Application (App)] An executable program. \wss{needs a
    proper definition, with citation(s)}
\item[Library] Services that are available to other programs, but not an
  executable program itself. \wss{needs a
    proper definition, with citation(s)}
\item[Artifacts] Work products generated during the process of creating
  software.  The work products include requirements documentation, design
  documentation, verification and validation plans, verification and validation
  reports, contributor's guides, user guides, build scripts, code, test cases,
  etc.
\end{description}

@@

\begin{itemize}
\item The software project is going to be long lived, where long life means at
  least 10 years.
\item The software+artifacts should interest multiple stakeholders (not just the
  original developer), with different interests and backgrounds.
\item The software is safety relevant, such as software for nuclear safety,
  medical imaging or computational medicine.
\item The software explicitly, or implicitly, is part of a program family of
  related software.  When the program family assumption means that there are
  related programs that are less effort to design and build together than to
  design and build as separate projects.  Program family development depends on
  satisfying three hypotheses \citep{Weiss1997}:
\begin{itemize}
\item Redevelopment hypothesis – most software development involved in producing
  an individual family members should be redevelopment because the family
  members have so much in common.
\item Oracle Hypothesis – the changes that are likely to occur during the soft-
  ware’s lifetime are predictable.
\item Organizational Hypothesis – designers can organize the software and the
  development effort so that predicted changes can be made independently.
\end{itemize}
\item The project recognizes that software artifacts other than the code are
  relevant.
\item The domain of the software project is well understood.  This assumption is
  related to the oracle hypothesis.
\end{itemize}

@@

\begin{enumerate}
\item What knowledge is currently captured in the artifacts generated by
  existing open source SCS projects?
\item What principles, processes, methodologies and tools are used by existing
  open source SCS projects?
\item What are the ``pain points'' for developers working on SCS projects?  What
  aspects of the existing processes, methodologies and tools do they consider
  could potentially be improved?
\item For a given software product (software+artifacts), how can we produce a
  reasonable measure of the software qualities (correctness, reliability,
  usability etc.) with a few hours of effort?
\item For a given software product (software+artifacts), how can we produce a
  meaningful measure of usability with a few days worth of effort?
\item For a given software product (software+artifacts), how can we produce a
  meaningful measure of maintainability (with respect to likely changes) with a
  few days worth of effort?
\item For a given software product (software+artifacts), how can we produce a
  meaningful measure of reproducibility/replicability with a few days worth of
  effort?
\item With a reasonable effort and time, how can we compare the productivity
  between different processes, methodologies and tools?
\item Can a correlation be identified between software projects that use best
  practices for the development process and methodologies and improvement in the
  qualities of usability, performance and reproducibility/replicability?
\end{enumerate}

>>>>

In the above definition, software is considered to be an application (app)
and/or a library.  The outcome of the development process is not just software.
It is \{software + artifacts\}.  \wss{softwartifacts?} \ad{try softifacts?}The artifacts consist of
such work products as documentation, build scripts, code, test cases,
contributor's guides, etc.  The quality of sustainability cannot be assessed by
looking at the software alone, or the code alone; it depends on everything,
including the processes and methodologies used \ad{during the development?}.  Further information on the
relevant definitions and context is given in the next section.

The notion of ``reasonable amount of time and energy'' is part of the objective,
but reasonable is not defined in the objective.  What is reasonable is going to
depend on the context.  Problems that are large, complex and/or critical are
going to have a greater ``budget'' of time and energy.  When comparing different
alternatives for processes and methodologies, preference would be given to the
alternative that costs less time and energy, assuming that this alternative
achieves the same \ad{or higher} level of sustainability as its competition.  The assumptions
listed in the next section outline the typical space of problems where it makes
sense to aim for sustainable \{software + artifacts\}.

@@

\begin{description}
\item[Knowledge] Facts, definitions, theories, assumptions, etc. \wss{needs a
    proper definition, with citation(s)}

``Awareness or familiarity gained by experience of a fact or situation.'' \citep{OxfordKnowledge2020}
\ad{Definition from Lexico.com (Oxford) dictionary}

\item[Process] Organization of the software development activities.  \wss{needs a
    proper definition, with citation(s)}

``A series of actions or steps taken in order to achieve a particular end.''\citep{OxfordProcess2020}
\ad{Definition from Lexico.com (Oxford) dictionary}

\item[Principles] Principles are ``general and abstract statements describing
  desirable properties of software processes and products.''  \citep[p.\
  41]{GhezziEtAl2003}
\item[Methods] ``Methods are general guidelines that govern the execution of some activity;
  they are rigorous, systematic, and disciplined approaches.''  \citep[p.\
  41]{GhezziEtAl2003}
\item[Techniques] ``Techniques are more technical and mechanical than methods.''
  \cite[p.\ 41]{GhezziEtAl2003}
\item[Methodology] Combination of methods and techniques.  \citep[p.\
  41]{GhezziEtAl2003}
\item[Tools] ``Tools ... are developed to support the application of techniques,
  method and methodologies.'' \cite[p.\ 41]{GhezziEtAl2003}

@@

This definition has brought in several other qualities.  These qualities are
defined in the ``Quality  Definitions of Qualities'' document.
\wss{Reproducibility above might be changed to replicability, or possibly both
  terms should be included?}
\wss{Should performance be included in the list of qualities?}
\ad{Almost every SCS needs to be correct, reliable and usable, but many of them
may not need to provide edge performance. How do we choose which qualities to be
included here?}

\item[Software] Application and/or library.
\item[Application (App)] An executable program. \wss{needs a
    proper definition, with citation(s)}

``Application software is a program or group of programs designed for end users.
These programs are divided into two classes: system software and application
software. While system software consists of low-level programs that interact
with computers at a basic level, application software resides above system
software and includes applications such as database programs, word processors
and spreadsheets. Application software may be bundled with system software or
published alone.''\citep{TechopediaApplication2018}\ad{Definition from
Techopedia.com}

``A program or piece of software designed and written to fulfill a particular
purpose of the user.''\citep{OxfordApplication2020}\ad{Definition from
Lexico.com (Oxford) dictionary}

\item[Library] Services that are available to other programs, but not an
  executable program itself. \wss{needs a
    proper definition, with citation(s)}

``A software library is a suite of data and programming code that is used to
develop software programs and applications. It is designed to assist both the
programmer and the programming language compiler in building and executing
software.''\citep{TechopediaLibrary2016}\ad{Definition from Techopedia.com}

``A collection of programs and software packages made generally available, often
loaded and stored on disk for immediate
use.''\citep{OxfordLibrary2020}\ad{Definition from Lexico.com (Oxford)
dictionary}

\item[Artifacts] Work products generated during the process of creating
  software.  The work products include requirements documentation, design
  documentation, verification and validation plans, verification and validation
  reports, contributor's guides, user guides, build scripts, code, test cases,
  etc.
\end{description}

@@

\begin{itemize}
\item The software project is going to be long lived, where long life means at
  least 10 years.
\item The software+artifacts should interest multiple stakeholders (not just the
  original developer), with different interests and backgrounds.
\item The software is safety relevant, such as software for nuclear safety,
  medical imaging or computational medicine.
\item The software explicitly, or implicitly, is part of a program family of
  related software.  When the program family assumption means that there are
  related programs that are less effort to design and build together than to
  design and build as separate projects.  Program family development depends on
  satisfying three hypotheses \citep{Weiss1997}:
\begin{itemize}
\item Redevelopment hypothesis – most software development involved in producing
  an individual family members should be redevelopment because the family
  members have so much in common.
\item Oracle Hypothesis – the changes that are likely to occur during the soft-
  ware’s \ad{software's} lifetime are predictable.
\item Organizational Hypothesis – designers can organize the software and the
  development effort so that predicted changes can be made independently.
\end{itemize}
\item The project recognizes that software artifacts other than the code are
  relevant.
\item The domain of the software project is well understood.  This assumption is
  related to the oracle hypothesis.
\end{itemize}

@@

\begin{enumerate}
\item What knowledge is currently captured in the artifacts generated by
  existing open source SCS projects?
\item What principles, processes, methodologies and tools are used by existing
  open source SCS projects?
\item What are the ``pain points'' for developers working on SCS projects?  What
  aspects of the existing processes, methodologies and tools do they consider
  could potentially be improved?
\item For a given software product (software+artifacts), how can we produce a
  reasonable measure of the software qualities (correctness, reliability,
  usability etc.) with a few hours of effort?
\item For a given software product (software+artifacts), how can we produce a
  meaningful measure of usability with a few days worth of effort?
\item For a given software product (software+artifacts), how can we produce a
  meaningful measure of maintainability (with respect to likely changes) with a
  few days worth of effort?
\item For a given software product (software+artifacts), how can we produce a
  meaningful measure of reproducibility/replicability with a few days worth of
  effort?
\item With a reasonable effort and time, how can we compare the productivity
  between different processes, methodologies and tools?
\item Can a correlation be identified between software projects that use best \ad{better}
  practices for the development process and methodologies and improvement in the
  qualities of usability, performance and reproducibility/replicability?
\ad{Change "best" to "better"? The best one isn't likely to be found, and better
practices may also result in improvements.}
\end{enumerate}
